---
title: 'Reproducible Research: Peer Assessment 1'
author: "Kavya Mohandas"
date: "24/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

# Calculate the total steps taken per day
1. Load the data . 
```{r}
activityData <- read.csv(file="activity.csv", header=TRUE)
totalSteps <- aggregate(steps ~ date, activityData, FUN=sum)
```



## What is mean total number of steps taken per day?
2. Make a histogram of the total number of steps taken per day
```{r}
hist(totalSteps$steps,
     main = "Total Steps per Day",
     xlab = "Number of Steps")
```

3.Calculate and report the mean and median of total steps taken per day
```{r}
meanSteps <- mean(totalSteps$steps, na.rm = TRUE)
medSteps <- median(totalSteps$steps, na.rm = TRUE)
```

## What is the average daily activity pattern?

4.Make a time-series plot of the 5-minute interval and the average number of
steps taken, averaged acoss all days.
```{r}
library(ggplot2)
meanStepsByInt <- aggregate(steps ~ interval, activityData, mean)
ggplot(data = meanStepsByInt, aes(x = interval, y = steps)) +
  geom_line() +
  ggtitle("Average Daily Activity Pattern") +
  xlab("5-minute Interval") +
  ylab("Average Number of Steps") +
  theme(plot.title = element_text(hjust = 0.5))
```


## Imputing missing values

5.Calculate and report the total number of missing values in the dataset
```{r}
missingVals <- is.na(activityData$steps)
```


6. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
imp_activityData <- transform(activityData,
                              steps = ifelse(is.na(activityData$steps),
                                             meanStepsByInt$steps[match(activityData$interval, 
                                                                        meanStepsByInt$interval)],
                                             activityData$steps))

```

7.Make a histogram of the total number of steps taken each day and and report the mean and median.
```{r}
impStepsByInt <- aggregate(steps ~ date, imp_activityData, FUN=sum)
hist(impStepsByInt$steps,
     main = "Imputed Number of Steps Per Day",
     xlab = "Number of Steps")


impMeanSteps <- mean(impStepsByInt$steps, na.rm = TRUE)
impMedSteps <- median(impStepsByInt$steps, na.rm = TRUE)
diffMean = impMeanSteps - meanSteps
diffMed = impMedSteps - medSteps
diffTotal = sum(impStepsByInt$steps) - sum(totalSteps$steps)
```

## Are there differences in activity patterns between weekdays and weekends?

8. Create a new factor variable in the dataset with two levels - "weekend" and "weekday"
```{r}
DayType <- function(date) {
  day <- weekdays(date)
  if (day %in% c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday'))
    return ("weekeday")
  else if (day %in% c('Saturday', 'Sunday'))
    return ("weekend")
  else
    stop ("Invalid Date Format.")
}
imp_activityData$date <- as.Date(imp_activityData$date)
imp_activityData$day <- sapply(imp_activityData$date, FUN = DayType)
```

9. Make a panel plot containnig a time-series plot of the 5-minute interval and the average number of steps taken across all weekdays or weekends
```{r}
meanStepsByDay <- aggregate(steps ~ interval + day, imp_activityData, mean)
ggplot(data = meanStepsByDay, aes(x = interval, y = steps)) + 
  geom_line() +
  facet_grid(day ~ .) +
  ggtitle("Average Daily Activity Pattern") +
  xlab("5-minute Interval") +
  ylab("Average Number of Steps") +
  theme(plot.title = element_text(hjust = 0.5))
```






